A Closer Look at Network Flow on Evolving Graphs

### 2018-10-10 paper.tex
update experiments and conclusion part

In the 1985 science-fiction film ``Back to the Future'', Marty McFly (played by Michael J. Fox) travelled back in time and accidentally changed history.
However, we cannot travel back in time or even send a message to the past.
To traverse an evolving graph, a.k.a a time-dependent graph, one has to consider order of time stamps. In particular, a walk on an evolving graph needs to be time-preserving, meaning we
can not pass a massage (through an edge) to a node at a earlier time.
Time-preversing walks and paths are very important concepts for
network flow on evolving graphs.

% time-respecting walk
We represent an evolving graph as a time-ordered sequences of graphs, similar to the work of Tang and coworkers \cite{nicosia13, tang09, tang102, tang10} and Grindrod, Higham and coworkers \cite{grindrod13,grindrod11}. The idea is to divide the system's time span into temporal slices and regards it as a sequence of static graphs $G^{[t]}$, one for each layer. Such representation arises naturally in many real-world applications. For example,
consider networks of users interacting through messaging. Each interaction between two users $A$ and $B$ are stored with a time stamp $t$. If we represent a user as a node then a user $A$ at time stamp $t$ is represented by a pair $(v,t)$ and a message sent from user $A$ to user $B$ at time stamp $t$ can be represented as $\langle (A, t) (B, t) \rangle$.

In general, a node $v$ on $G^{[t]}$ is represented by a pair $(v, t)$, where $t$ is the time stamp of the graph $G^{[t]}$.
A sequence of interactions between users form a walk. Suppose $A$ sent a message to $B$ at time stamp $t_1$
and then $B$ sent this message to $C$ at time stamp $t_2$. We can represent this activity as $\langle (A,t_1), (B,t_1), (B, t_2), (C,t_2)\rangle$, where each pair of two nodes has meanings: $\langle (A,t_1), (B,t_1) \rangle$ and $\langle (B,t_2), (C,t_2) \rangle$ represent interactions between users $A$, $B$, and $C$ and $\langle (B,t_1), (B,t_2) \rangle$ represents the fact that user $B$ holds the message at time stamp $t_1$ and $t_2$.

A time-respecting walk is a sequence of nodes $\langle (v_1, t_1), (v_2, t_2), \ldots ,(v_m, t_m)\rangle$, where $t_1 \le t_2 \le \cdots \le t_n$ are the corresponding time stamps. If $t_i = t_j$, then $\langle (v_i, t_i), (v_j, t_j) \rangle$ is an edge where both nodes are on the same temporal slice,  otherwise it is an edge linking nodes from different temporal slices.
We call edges of the first kind \emph{static edges} and of the second kind \emph{causal edges}.
Recall in static graph, a walk of length $m$ is defined as a sequence of nodes $v_1, v_2, \ldots, v_{m+1}$ such that for each $i = 1, \ldots, m$ there is an edge from $v_i$ to $v_{i+1}$. How to determine the length of a time-respecting walk?
One could define the length of a time-respecting walk as the total number of temporal slices (see shortest temporal distance in \cite{tang09}) or the number of static edges (see dynamic walk in \cite{grindrod11}). This paper studies the impact of these two different definitions. In particular, we studies a general definition of temporal distance that takes account of both temporal slices and static edges.

@@

% no work on pagerank so far. By understanding temporal walk, we could provide a generalization of pagerank on evolving graphs.

% In Higham's following up papers, time delay is considered as alternative format of Katz centrality.
% But the impact of temporal is left unconsidered. We studied the problem on centrality from two different types of centrality: one that depends on matrix unfolding and one that depends on temporal walk.

% our contribution this paper
Borgatti \cite{borgatti05} noted that the manner of traffic flow entails a new way to think about centrality. In particularly, the importance of a node in a network can not be determined without reference to how traffic flows through the network.
Inspired by this idea, we simulate network flows in evolving graphs to measure the centrality scores.
% Grindrod et al.'s definition of dynamic walk \cite{grindrod11} only counts static edges.
% Tang et al. \cite{tang10s} measure a temporal shortest walk in time unit (or causal edges) only.
% If a node can be reached by another node in $k$ time stamps then the distance between the two nodes is $k$. Chen and Zhang \cite{chen16} considered both static edges and causal edges in their definition of temporal walk.
The aims of this paper are as follows. First, to derive important evolving graph centrality from first principle via network flow. Second, to study the impact of temporal distance between nodes on centrality.
Third, to generalise PageRank on evolving graphs.
We carry out the experiment using the Julia dynamic network analysis package
EvolvingGraphs.jl (\url{https://github.com/EtymoIO/EvolvingGraphs.jl}).
We use the research paper data gathered from Etymo (\url{https://etymo.io}), a visual search engine for data scientists.

@@

We let $G_n = \langle G^{[1]}, G^{[2]}, \ldots, G^{[n]} \rangle$ be an evolving graph
and $A^{[1]}, A^{[2]}, \dots A^{[n]}$ be the corresponding adjacency matrices.
Then the dynamic communicability matrix is defined as
\begin{equation}
  \label{eq:q}
  Q^{[k]} = (I - \alpha A^{[1]})^{-1}(I - \alpha A^{[2]})^{-1} \cdots (I -  \alpha A^{[n]})^{-1},
\end{equation}
where parameter $\alpha < 1/ \max_k \rho(A^{[k]})$, the spectral radius of $A^{[k]}$.
The $i$th row sum of $Q^{[k]}$ is the broadcast centrality.

Suppose $A$, $B$, and $C$ are three co-workers in a large internet company. Figure \ref{fig:katz_eg}
shows how they exchange ideas during $3$ days at work. The directions of the arrows indicate the direction of the flow of ideas.
For example, $A$ shares a new idea with $B$ on day $1$ and $B$ shares a new idea with $C$ on dat $3$. As a result, $A$'s idea can influence $C$.

\begin{figure}[h]
 \begin{center}
    \begin{tikzpicture}[scale=.38, line width =1.5pt]
  \node[circle,fill=green1, minimum size=0.2cm] (n7) at (-5,7) {B};
  \node[circle,draw=red1, minimum size=0.2cm] (n8) at (-10,5) {C};
  \node[circle,fill=green1, minimum size=0.2cm] (n10) at (-8,10) {A};
  \node[circle,fill=green1, minimum size=0.2cm] (n6) at (3,7) {B};
  \node[circle,fill=green1, minimum size=0.2cm] (n4) at (-2,5) {C};
  \node[circle,fill=green1, minimum size=0.2cm] (n1) at (0,10) {A};
   \node[circle,fill=green1, minimum size=0.2cm] (n11) at (11,7) {B};
  \node[circle,fill=green1, minimum size=0.2cm] (n12) at (6,5)  {C};
  \node[circle,draw=red1, minimum size=0.2cm] (n14) at (8,10) {A};
  \foreach \from/\to in {n10/n7, n1/n4, n4/n1, n11/n12, n1/n6}
   \draw[every edge,->] (\from) -- (\to);
 %\draw[dotted,->](n4) -- (n12);
% \draw[dotted,->](n7) to[out=80, in=40](n11);
\draw [decorate,decoration={brace,amplitude=5pt},xshift=-4pt,yshift=0pt]
(-4,3) -- (-10.5,3) node [midway,yshift=-0.6cm]{ $1$};
\draw [decorate,decoration={brace,amplitude=5pt},xshift=-4pt,yshift=0pt]
(4,3) -- (-2,3) node [midway,yshift=-0.6cm]{$2$};
\draw [decorate,decoration={brace,amplitude=5pt},xshift=-4pt,yshift=0pt]
(11.5,3) -- (5.5,3) node [midway,yshift=-0.6cm]{ $3$};
    \end{tikzpicture}
\end{center}
\caption{An evolving directed graph with 3 time stamps $1$, $2$ and $3$.
At each time stamp, the evolving graph is represented as a graph.
The green filled circles represent active nodes while the red circles represent
inactive nodes. Directed edges in each time stamp are shown as black arrows.}
\label{fig:katz_eg}
\end{figure}

@@

As expected $A$ is the most important node in the network. $C$ is the second important node in the network as it influenced
$A$ at time stamp $2$. If we reverse the time of communication between day $1$ and $3$, i.e., $B$ shares a new idea with $C$ on day $1$ and $A$ shares a new idea with $B$ on day $3$. Then in this case the idea $A$ had on day $3$ can not pass to $C$. This is illustrated in Figure \ref{fig:katz_eg2}.

\begin{figure}[h]
 \begin{center}
    \begin{tikzpicture}[scale=.38, line width =1.5pt]
  \node[circle,fill=green1, minimum size=0.2cm] (n7) at (-5,7) {B};
  \node[circle,fill=green1, minimum size=0.2cm] (n8) at (-10,5) {C};
  \node[circle,draw=red1, minimum size=0.2cm] (n10) at (-8,10) {A};
  \node[circle,fill=green1, minimum size=0.2cm] (n6) at (3,7) {B};
  \node[circle,fill=green1, minimum size=0.2cm] (n4) at (-2,5) {C};
  \node[circle,fill=green1, minimum size=0.2cm] (n1) at (0,10) {A};
   \node[circle,fill=green1, minimum size=0.2cm] (n11) at (11,7) {B};
  \node[circle,draw=red1, minimum size=0.2cm] (n12) at (6,5)  {C};
  \node[circle,fill=green1, minimum size=0.2cm] (n14) at (8,10) {A};
  \foreach \from/\to in {n7/n8, n1/n4, n4/n1, n1/n6, n14/n11}
   \draw[every edge,->] (\from) -- (\to);
 %\draw[dotted,->](n4) -- (n12);
% \draw[dotted,->](n7) to[out=80, in=40](n11);
\draw [decorate,decoration={brace,amplitude=5pt},xshift=-4pt,yshift=0pt]
(-4,3) -- (-10.5,3) node [midway,yshift=-0.6cm]{ $1$};
\draw [decorate,decoration={brace,amplitude=5pt},xshift=-4pt,yshift=0pt]
(4,3) -- (-2,3) node [midway,yshift=-0.6cm]{ $2$};
\draw [decorate,decoration={brace,amplitude=5pt},xshift=-4pt,yshift=0pt]
(11.5,3) -- (5.5,3) node [midway,yshift=-0.6cm]{ $3$};
    \end{tikzpicture}
\end{center}
\caption{An evolving directed graph with 3 time stamps $1$, $2$ and $3$.
At each time stamp, the evolving graph is represented as a graph.
The green filled circles represent active nodes while the red circles represent
inactive nodes. Directed edges in each time stamp are shown as black arrows.}
\label{fig:katz_eg2}
\end{figure}

In this cast, we have
\begin{lstlisting}
g2 = EvolvingGraph{Node{String}, Int}()
add_bunch_of_edges!(g2, [("A", "B", 3), ("A", "C", 2),
("A", "B", 2),("C", "A", 2),("B", "C", 1)])
\end{lstlisting}

The centrality ratings of the three nodes are

\begin{lstlisting}
katz(g2)
3-element Array{Tuple{EvolvingGraphs.Node{String},Float64},1}:
 (Node(A), 0.687679)
 (Node(B), 0.490062)
 (Node(C), 0.535666)
\end{lstlisting}

Notice the rating of $A$ decreased as its influence is not as important as in Figure \ref{fig:katz_eg}.
% does not take account of walks in a long time. cost in time
If the communication between $A$, $B$, and $C$ are not on three consecutive days
but on day $1$, day $10$, and day $100$, as shown in Figure \ref{fig:katz_eg3},
we would expect the centrality scores to change. However, according to \eqref{eq:q}, scores are unchanged.
We notice that the generalised Katz centrality on evolving graphs only takes account of the edges at each time stamp but not the communication time between edges at different time stamps.

\begin{figure}[h]
 \begin{center}
    \begin{tikzpicture}[scale=.38, line width =1.5pt]
  \node[circle,fill=green1, minimum size=0.2cm] (n7) at (-5,7) {B};
  \node[circle,fill=green1, minimum size=0.2cm] (n8) at (-10,5) {C};
  \node[circle,draw=red1, minimum size=0.2cm] (n10) at (-8,10) {A};
  \node[circle,fill=green1, minimum size=0.2cm] (n6) at (3,7) {B};
  \node[circle,fill=green1, minimum size=0.2cm] (n4) at (-2,5) {C};
  \node[circle,fill=green1, minimum size=0.2cm] (n1) at (0,10) {A};
   \node[circle,fill=green1, minimum size=0.2cm] (n11) at (11,7) {B};
  \node[circle,draw=red1, minimum size=0.2cm] (n12) at (6,5)  {C};
  \node[circle,fill=green1, minimum size=0.2cm] (n14) at (8,10) {A};
  \foreach \from/\to in {n7/n8, n1/n4, n4/n1, n1/n6, n14/n11}
   \draw[every edge,->] (\from) -- (\to);
 %\draw[dotted,->](n4) -- (n12);
% \draw[dotted,->](n7) to[out=80, in=40](n11);
\draw [decorate,decoration={brace,amplitude=5pt},xshift=-4pt,yshift=0pt]
(-4,3) -- (-10.5,3) node [midway,yshift=-0.6cm]{ $1$};
\draw [decorate,decoration={brace,amplitude=5pt},xshift=-4pt,yshift=0pt]
(4,3) -- (-2,3) node [midway,yshift=-0.6cm]{ $10$};
\draw [decorate,decoration={brace,amplitude=5pt},xshift=-4pt,yshift=0pt]
(11.5,3) -- (5.5,3) node [midway,yshift=-0.6cm]{ $100$};
    \end{tikzpicture}
\end{center}
\caption{An evolving directed graph with 3 time stamps $1$, $10$ and $100$.
At each time stamp, the evolving graph is represented as a graph.
The green filled circles represent active nodes while the red circles represent
inactive nodes. Directed edges in each time stamp are shown as black arrows.}
\label{fig:katz_eg3}
\end{figure}

@@

% what is walk and what is path?
A walk of length $l$ is a sequence of nodes $v_1, v_2, \ldots, v_l, v_{l+1}$ such that for
each $i = 1, 2, \ldots, l$, there is an edge from $v_i$ to $v_{i+1}$. A path of length $l$ is a walk of length $l$
such that all the nodes are different.
% relate to centrality
The definition of centrality depends on the manner in which traffic flows through a network \cite{borgatti05}.
Time-preserving walks induce temporal Katz centrality and temporal closeness centrality while time-preserving paths induce betweenness centrality and closeness centrality.

@@

\begin{definition}
  An \textbf{evolving graph} $G_n$ is a sequence of (static) graphs
$G_n = \langle G^{[1]}, G^{[2]},  \ldots ,G^{[n]} \rangle$ with associated time stamps
$t_1, t_2, \ldots, t_n$ respectively. Each $G^{[t]} = (V^{[t]}, E^{[t]})$ represents a (static) graph labelled by a time t.
\end{definition}

Note the node sets $V^{[t]}$ can change over time, i.e., nodes may appear or disappear at a particular time stamp.
For example, in Figure \ref{fig:eg_shortest_walk}, at time stamp $1998$, $V^{[1998]} = \langle 1, 2 \rangle$ and $E^{[1998]} = \langle (1,2) \rangle$. Each graph $G^{[t]}$ can be represented by its adjacency matrix $A^{[t]}$.
We can represent $G_n$ by a list of adjacency matrices $A_n = \langle A^{[1]}, A^{[2]}, \ldots, A^{[n]} \rangle$.

If we only measure the temporal difference in a temporal walk, we have a new definition of Katz centrality.
In the matrix form, the nonzero entries of $A^{[t_i]} A^{[t_{i+1}]}$ counts all the temporal walks of length $1$, and
$A^{[t_i]}\cdots A^{[t_j]}$, where $i < j$ counts all the temporal walks of length $j -i$.
For example, $A^{[1]}A^{[4]}$, $A^{[1]}A^{[1]}A^{[4]}$ and $A^{[1]}A^{[2]}A^{[3]}A^{[4]}$ are all walks of length 3.
In the original Katz centrality, the `attenuation' factor $\alpha$ is imposed on the
number of paths, i.e., longer paths has smaller effect to the overall rating.
Hence the final result is the summation of all products of the form
\[
\alpha^k A^{[i]} \cdots A^{[i+k]}
\]
Note this can not be written in \eqref{eq:q} as discussed in Section \ref{sec:temp-katz-centr}.

@@

In Figure \ref{fig:eg_shortest_walk}, the filled green circles are active nodes. For the adjacency matrix representation, a node $i$ is active at time stamp $t$, then
the $i$th row or the $i$th column of $A^{[t]}$ has at least one none-zero entry.

@@

 For example, in Figure \ref{fig:eg_shortest_walk}
 $\langle (1, 1998), (2, 1998) \rangle$ is a static edge at time stamp $1998$ and $\langle (3, 1999), (3, 2018) \rangle$ is a causal edge between time stamp $1999$ and time stamp $2018$.
 The difference between static edges and causal edges were first explicitly considered in \cite{chen16}.

 We can denote a \emph{weighted static edge} as $\langle (v_i, t), (v_j, t), w_s \rangle$ and a \emph{weighted causal edge} as $\langle (v, t_i), (v, t_j), w_t \rangle$.
 Here $w_s$ represents the spatial distance between the two nodes $v_i$ and $v_j$; $w_t$ represents the
 temporal distance of $v$ at different time stamps.
 In Figure \ref{fig:eg_shortest_walk}, the black arrows represent static edges and the dotted arrows represent causal edges.
 Suppose we assign all the static edge weight to $1$ and
 let the causal edge weight be the number of time stamps between the pair of nodes. Then in Figure \ref{fig:eg_shortest_walk}
 the edge weight between $(1, 1998)$ and $(2, 1998)$ is $1$ and the edge weight between $(2, 1998)$ and $(2, 2018)$ is $10$.
In general, a time-preserving walk can be defined as follows.

\begin{definition}
A \textbf{time-preserving walk}  of length $m$ on an evolving graph $G_n$ is a time-ordered sequence of active nodes, $\langle (v_1, t_1), (v_2, t_2), \ldots, (v_m, t_m) \rangle$, where $t_1 \le t_2 \le \cdots \le t_m$ and
$v_i = v_j$ if and only if $t_i \ne t_j$.
\end{definition}

>>>>

In the 1985 science-fiction film ``Back to the Future'', Marty McFly (played by Michael J. Fox) travelled back in time and accidentally changed history.
However, we cannot travel back in time or even send a message to the past.
To traverse an evolving graph (a time-dependent graph), one has to consider the order of time stamps. In particular, a walk on an evolving graph needs to be time-preserving, meaning we
can not pass a massage (through an edge) to a node at a earlier time.
Time-preversing walks and paths are very important concepts for
network flow on evolving graphs.

% time-respecting walk
We represent an evolving graph as a time-ordered sequences of graphs, similar to the work of Tang and coworkers \cite{ntmm13, tmml09, tmml10, tsmm09} and Grindrod, Higham and coworkers \cite{grihig13,gphe11}. The idea is to divide the system's time span into temporal slices and regard it as a sequence of static graphs $G^{[t]}$, one for each layer. Such a representation arises naturally in many real-world applications. For example,
consider networks of users interacting through messaging. Each interaction between two users $A$ and $B$ is stored with a time stamp $t$. If we represent user $A$ as a node then user $A$ at time stamp $t$ is represented by a pair $(A,t)$ and a message sent from user $A$ to user $B$ at time stamp $t$ can be represented as $\langle (A, t), (B, t) \rangle$.

In general, a node $v$ on $G^{[t]}$ is represented by a pair $(v, t)$, where $t$ is the time stamp of the graph $G^{[t]}$.
A sequence of interactions between users form a walk. Suppose $A$ sent a message to $B$ at time stamp $t_1$
and then $B$ sent this message to $C$ at time stamp $t_2$. We can represent this activity as $\langle (A,t_1), (B,t_1), (B, t_2), (C,t_2)\rangle$, where each pair of two nodes has meanings: $\langle (A,t_1), (B,t_1) \rangle$ and $\langle (B,t_2), (C,t_2) \rangle$ represent interactions between users $A$, $B$, and $C$ and $\langle (B,t_1), (B,t_2) \rangle$ represents the fact that user $B$ holds the message at time stamps $t_1$ and $t_2$.

A time-respecting walk is a sequence of nodes $\langle (v_1, t_1), (v_2, t_2), \ldots ,(v_m, t_m)\rangle$, where $t_1 \le t_2 \le \cdots \le t_n$ are the corresponding time stamps. If $t_i = t_j$, then $\langle (v_i, t_i), (v_j, t_j) \rangle$ is an edge where both nodes are on the same temporal slice,  otherwise it is an edge linking nodes from different temporal slices.
We call edges of the first kind \emph{static edges} and of the second kind \emph{causal edges}.
Recall that in static graph, a walk of length $m$ is defined as a sequence of nodes $v_1, v_2, \ldots, v_{m+1}$ such that for each $i = 1, \ldots, m$ there is an edge from $v_i$ to $v_{i+1}$. How can we  determine the length of a time-respecting walk?
One could define the length of a time-respecting walk as the total number of temporal slices (see the shortest temporal distance in \cite{tmml09}) or the number of static edges (see the dynamic walk in \cite{gphe11}). This paper studies the impact of these two different definitions. In particular, we study a general definition of temporal distance that takes account of both temporal slices and static edges.

@@

% our contribution this paper
Borgatti \cite{borgatti05} noted that the manner of traffic flow entails a new way to think about centrality. In particular, the importance of a node in a network cannot be determined without reference to how traffic flows through the network.
Inspired by this idea, we simulate network flows in evolving graphs to measure the centrality scores.
% Grindrod et al.'s definition of dynamic walk \cite{gphe11} only counts static edges.
% Tang et al. \cite{tang10s} measure a temporal shortest walk in time unit (or causal edges) only.
% If a node can be reached by another node in $k$ time stamps then the distance between the two nodes is $k$. Chen and Zhang \cite{chen16} considered both static edges and causal edges in their definition of temporal walk.
The aims of this paper are as follows. First, to derive important evolving graph centrality from first principles via network flow. Second, to study the impact of temporal distance between nodes on centrality.
Third, to generalise PageRank on evolving graphs.
We carry out the experiment using the Julia dynamic network analysis package
EvolvingGraphs.jl (\url{https://github.com/EtymoIO/EvolvingGraphs.jl}).
We use the research paper data gathered from Etymo (\url{https://etymo.io}), a visual search engine for data scientists.

@@

We let $G_n = \langle G^{[1]}, G^{[2]}, \ldots, G^{[n]} \rangle$ be an evolving graph
and $A^{[1]}, A^{[2]}, \dots A^{[n]}$ be the corresponding adjacency matrices.
Then the dynamic communicability matrix is defined as
\begin{equation}
  \label{eq:q}
  Q^{[k]} = (I - \alpha A^{[1]})^{-1}(I - \alpha A^{[2]})^{-1} \cdots (I -  \alpha A^{[n]})^{-1},
\end{equation}
where the parameter $\alpha$ satisfies $\alpha < 1/ \max_k \rho(A^{[k]})$, the spectral radius of $A^{[k]}$.
The $i$th row sum of $Q^{[k]}$ is the broadcast centrality.

Suppose $A$, $B$, and $C$ are three co-workers in a large internet company. Figure \ref{fig:katz_eg}
shows how they exchange ideas during three days at work. The directions of the arrows indicate the direction of the flow of ideas.
For example, $A$ shares a new idea with $B$ on day $1$ and $B$ shares a new idea with $C$ on dat $3$. As a result, $A$'s idea can influence $C$. Here, and in subsequent figures, green filled circles represent active node and red circles represent inactive nodes.

\begin{figure}[h]
 \begin{center}
    \begin{tikzpicture}[scale=.38, line width =1.5pt]
  \node[circle,fill=green1, minimum size=0.2cm] (n7) at (-5,7) {B};
  \node[circle,draw=red1, minimum size=0.2cm] (n8) at (-10,5) {C};
  \node[circle,fill=green1, minimum size=0.2cm] (n10) at (-8,10) {A};
  \node[circle,fill=green1, minimum size=0.2cm] (n6) at (3,7) {B};
  \node[circle,fill=green1, minimum size=0.2cm] (n4) at (-2,5) {C};
  \node[circle,fill=green1, minimum size=0.2cm] (n1) at (0,10) {A};
   \node[circle,fill=green1, minimum size=0.2cm] (n11) at (11,7) {B};
  \node[circle,fill=green1, minimum size=0.2cm] (n12) at (6,5)  {C};
  \node[circle,draw=red1, minimum size=0.2cm] (n14) at (8,10) {A};
  \foreach \from/\to in {n10/n7, n1/n4, n4/n1, n11/n12, n1/n6}
   \draw[every edge,->] (\from) -- (\to);
 %\draw[dotted,->](n4) -- (n12);
% \draw[dotted,->](n7) to[out=80, in=40](n11);
\draw [decorate,decoration={brace,amplitude=5pt},xshift=-4pt,yshift=0pt]
(-4,3) -- (-10.5,3) node [midway,yshift=-0.6cm]{ $1$};
\draw [decorate,decoration={brace,amplitude=5pt},xshift=-4pt,yshift=0pt]
(4,3) -- (-2,3) node [midway,yshift=-0.6cm]{$2$};
\draw [decorate,decoration={brace,amplitude=5pt},xshift=-4pt,yshift=0pt]
(11.5,3) -- (5.5,3) node [midway,yshift=-0.6cm]{ $3$};
    \end{tikzpicture}
\end{center}
\caption[An evolving directed graph with 3 time stamps $1$, $2$ and $3$.]{An evolving directed graph with 3 time stamps $1$, $2$ and $3$.
At each time stamp, the evolving graph is represented as a graph.
The green filled circles represent active nodes while the red circles represent
inactive nodes. Directed edges in each time stamp are shown as black arrows.}
\label{fig:katz_eg}
\end{figure}

@@

As expected $A$ is the most important node in the network. $C$ is the second most important node in the network as it influenced
$A$ at time stamp $2$. If we reverse the time of communication between day $1$ and $3$, i.e., $B$ shares a new idea with $C$ on day $1$ and $A$ shares a new idea with $B$ on day $3$. Then in this case the idea $A$ had on day $3$ cannot pass to $C$. This is illustrated in Figure \ref{fig:katz_eg2}.
\begin{figure}[h]
 \begin{center}
    \begin{tikzpicture}[scale=.38, line width =1.5pt]
  \node[circle,fill=green1, minimum size=0.2cm] (n7) at (-5,7) {B};
  \node[circle,fill=green1, minimum size=0.2cm] (n8) at (-10,5) {C};
  \node[circle,draw=red1, minimum size=0.2cm] (n10) at (-8,10) {A};
  \node[circle,fill=green1, minimum size=0.2cm] (n6) at (3,7) {B};
  \node[circle,fill=green1, minimum size=0.2cm] (n4) at (-2,5) {C};
  \node[circle,fill=green1, minimum size=0.2cm] (n1) at (0,10) {A};
   \node[circle,fill=green1, minimum size=0.2cm] (n11) at (11,7) {B};
  \node[circle,draw=red1, minimum size=0.2cm] (n12) at (6,5)  {C};
  \node[circle,fill=green1, minimum size=0.2cm] (n14) at (8,10) {A};
  \foreach \from/\to in {n7/n8, n1/n4, n4/n1, n1/n6, n14/n11}
   \draw[every edge,->] (\from) -- (\to);
 %\draw[dotted,->](n4) -- (n12);
% \draw[dotted,->](n7) to[out=80, in=40](n11);
\draw [decorate,decoration={brace,amplitude=5pt},xshift=-4pt,yshift=0pt]
(-4,3) -- (-10.5,3) node [midway,yshift=-0.6cm]{ $1$};
\draw [decorate,decoration={brace,amplitude=5pt},xshift=-4pt,yshift=0pt]
(4,3) -- (-2,3) node [midway,yshift=-0.6cm]{ $2$};
\draw [decorate,decoration={brace,amplitude=5pt},xshift=-4pt,yshift=0pt]
(11.5,3) -- (5.5,3) node [midway,yshift=-0.6cm]{ $3$};
    \end{tikzpicture}
\end{center}
\caption{An evolving directed graph with 3 time stamps $1$, $2$ and $3$.}
\label{fig:katz_eg2}
\end{figure}
In this case, we have
\begin{lstlisting}
g2 = EvolvingGraph{Node{String}, Int}()
add_bunch_of_edges!(g2, [("A", "B", 3), ("A", "C", 2),
("A", "B", 2),("C", "A", 2),("B", "C", 1)])
\end{lstlisting}

The centrality ratings of the three nodes are
\begin{lstlisting}
katz(g2)
3-element Array{Tuple{EvolvingGraphs.Node{String},Float64},1}:
 (Node(A), 0.687679)
 (Node(B), 0.490062)
 (Node(C), 0.535666)
\end{lstlisting}

Notice the rating of $A$ decreased as its influence is not as important as in Figure \ref{fig:katz_eg}.
% does not take account of walks in a long time. cost in time
If the communication between $A$, $B$, and $C$ are not on three consecutive days
but on day $1$, day $10$, and day $100$, as shown in Figure \ref{fig:katz_eg3},
we would expect the centrality scores to change. However, according to \eqref{eq:q}, the scores are unchanged.
We notice that the generalised Katz centrality on evolving graphs only takes account of the edges at each time stamp but not the communication time between edges at different time stamps.

\begin{figure}[h]
 \begin{center}
    \begin{tikzpicture}[scale=.38, line width =1.5pt]
  \node[circle,fill=green1, minimum size=0.2cm] (n7) at (-5,7) {B};
  \node[circle,fill=green1, minimum size=0.2cm] (n8) at (-10,5) {C};
  \node[circle,draw=red1, minimum size=0.2cm] (n10) at (-8,10) {A};
  \node[circle,fill=green1, minimum size=0.2cm] (n6) at (3,7) {B};
  \node[circle,fill=green1, minimum size=0.2cm] (n4) at (-2,5) {C};
  \node[circle,fill=green1, minimum size=0.2cm] (n1) at (0,10) {A};
   \node[circle,fill=green1, minimum size=0.2cm] (n11) at (11,7) {B};
  \node[circle,draw=red1, minimum size=0.2cm] (n12) at (6,5)  {C};
  \node[circle,fill=green1, minimum size=0.2cm] (n14) at (8,10) {A};
  \foreach \from/\to in {n7/n8, n1/n4, n4/n1, n1/n6, n14/n11}
   \draw[every edge,->] (\from) -- (\to);
 %\draw[dotted,->](n4) -- (n12);
% \draw[dotted,->](n7) to[out=80, in=40](n11);
\draw [decorate,decoration={brace,amplitude=5pt},xshift=-4pt,yshift=0pt]
(-4,3) -- (-10.5,3) node [midway,yshift=-0.6cm]{ $1$};
\draw [decorate,decoration={brace,amplitude=5pt},xshift=-4pt,yshift=0pt]
(4,3) -- (-2,3) node [midway,yshift=-0.6cm]{ $10$};
\draw [decorate,decoration={brace,amplitude=5pt},xshift=-4pt,yshift=0pt]
(11.5,3) -- (5.5,3) node [midway,yshift=-0.6cm]{ $100$};
    \end{tikzpicture}
\end{center}
\caption[An evolving directed graph with 3 time stamps $1$, $10$ and $100$.]{An evolving directed graph with 3 time stamps $1$, $10$ and $100$.
At each time stamp, the evolving graph is represented as a graph.
The green filled circles represent active nodes while the red circles represent
inactive nodes. Directed edges in each time stamp are shown as black arrows.}
\label{fig:katz_eg3}
\end{figure}

@@

% what is walk and what is path?
A walk of length $\ell$ is a sequence of nodes $v_1, v_2, \ldots, v_l, v_{l+1}$ such that for
each $i = 1, 2, \ldots, \ell$, there is an edge from $v_i$ to $v_{i+1}$. A path of length $\ell$ is a walk of length $\ell$
such that all the nodes are different.
% what is time-preserving walks
Similarly, we define a temporal walk to be a time-ordered sequence of temporal nodes. A temporal path is a temporal walk such that all the temporal nodes are different.
% relate to centrality
The definition of centrality depends on the manner in which traffic flows through a network \cite{borgatti05}.
Time-preserving walks induce temporal Katz centrality and temporal closeness centrality while time-preserving paths induce betweenness centrality and closeness centrality.

@@

\begin{definition}
  An \textbf{evolving graph} $G_n$ is a sequence of (static) graphs
$G_n = \langle G^{[1]}, G^{[2]},  \ldots ,G^{[n]} \rangle$ with associated time stamps
$t_1 \leq t_2 \leq \ldots \leq t_n$.
Each $G^{[t]} = (V^{[t]}, E^{[t]})$ represents a (static) graph labelled by a time t.
\end{definition}

Note that the node sets $V^{[t]}$ can change over time, i.e., nodes may appear or disappear at a particular time stamp.
For example, in Figure \ref{fig:eg_shortest_walk}, at time stamp $1998$, $V^{[1998]} = \langle 1, 2 \rangle$ and $E^{[1998]} = \langle (1,2) \rangle$. Each graph $G^{[t]}$ can be represented by its adjacency matrix $A^{[t]}$.
We can represent $G_n$ by a list of adjacency matrices $A_n = \langle A^{[1]}, A^{[2]}, \ldots, A^{[n]} \rangle$.

If we only measure the temporal difference in a temporal walk, we have a new definition of Katz centrality.
In the matrix form, the nonzero entries of $A^{[t_i]} A^{[t_{i+1}]}$ counts all the temporal walks of length $1$, and
$A^{[t_i]}\cdots A^{[t_j]}$, where $i < j$ counts all the temporal walks of length $j -i$.
For example, $A^{[1]}A^{[4]}$, $A^{[1]}A^{[1]}A^{[4]}$ and $A^{[1]}A^{[2]}A^{[3]}A^{[4]}$ count all walks of length 3.
In the original Katz centrality, the `attenuation' factor $\alpha$ is imposed on the
number of paths, i.e., longer paths has smaller effect to the overall rating.
Hence the final result is the summation of all products of the form
\[
\alpha^k A^{[i]} \cdots A^{[i+k]}
\]
Note this is different from \eqref{eq:q} as discussed in Section \ref{sec:temp-katz-centr}.

@@

In Figure \ref{fig:eg_shortest_walk}, the filled green circles are active nodes. For the adjacency matrix representation, if a node $i$ is active at time stamp $t$ then
the $i$th row or the $i$th column of $A^{[t]}$ has at least one none-zero entry.

@@

 For example, in Figure \ref{fig:eg_shortest_walk}
 $\langle (1, 1998), (2, 1998) \rangle$ is a static edge at time stamp $1998$ and $\langle (3, 1999), (3, 2018) \rangle$ is a causal edge between time stamp $1999$ and time stamp $2018$.
 The difference between static edges and causal edges was first explicitly considered in \cite{chen16}.

 We can denote a \emph{weighted static edge} as $\langle (v_i, t), (v_j, t), w_s \rangle$ and a \emph{weighted causal edge} as $\langle (v, t_i), (v, t_j), w_t \rangle$.
 Here $w_s$ represents the spatial distance between the two nodes $v_i$ and $v_j$; $w_t$ represents the
 temporal distance of $v$ at different time stamps.
 In Figure \ref{fig:eg_shortest_walk}, the black arrows represent static edges and the dotted arrows represent causal edges.
 Suppose we assign all the static edge weights to $1$ and
 let the causal edge weight be the number of time stamps between the pair of nodes. Then in Figure \ref{fig:eg_shortest_walk}
 the edge weight between $(1, 1998)$ and $(2, 1998)$ is $1$ and the edge weight between $(2, 1998)$ and $(2, 2018)$ is $10$.
In general, a weighted time-preserving walk can be defined as follows.

\begin{definition}
A \textbf{weighted time-preserving walk} of length $m$ on an evolving graph $G_n$ is an alternating
time-ordered sequence of active nodes and edges, $\langle (v_1, t_1), e_1, (v_2, t_2), e_2, (v_3, t_3) \ldots, e_m, (v_m, t_m) \rangle$, where $t_1 \le t_2 \le \cdots \le t_m$ and
$v_i = v_j$ if and only if $t_i \ne t_j$, and $e_i$ can be either a weighted static edge or weighted causal edge.
\end{definition}
